Refinement of Thermostated Molecular Dynamics via Backward Error Analysis

### 2018-06-18 rigid_bodies.tex

In the previous section, we have seen that drifts in the extended energy of an NVT simulation can accumulate in the thermostat-related variables, thus preventing full-fledged instability from happening.
This allows simulations to be executed with large time steps, which could in principle save computer time.
However, as we will show now, discretization errors can become manifest much earlier than any risk of instability.
Fortunately, our proposed methodology is capable of mitigating the effect of these errors on the ensemble averages computed for the simulated system.

In Fig.~\ref{fig:properties}, we present some thermodynamic properties computed for our liquid water system using several NVT-MD methods with different time step sizes.
All simulations were carried out by specifying $T = 298~\text{K}$ and the analyzed quantities are the estimated temperature $T_\text{est} = \frac{2\langle K \rangle}{(6N-3)k_\mini{B}}$,
the mean potential energy per molecule $\langle U/N \rangle$,
the mean internal virial per molecule $\langle W/N \rangle$, and
the constant-volume specific heat $c_\mini{V} = \frac{\langle \Ham{}^2 \rangle - \langle \Ham{} \rangle^2}{N k_\mini{B} T^2}$.
The internal virial $W$ is computed as described in Ref.~\citenum{Silveira_2017}.
All symbols in Fig.~\ref{fig:properties} represent results obtained after $18~\text{ns}$ of production time, while the lines are just guides for the eyes.
The horizontal, dotted lines are meant to help visualizing the departure of the computed averages from their expected values.
They correspond to the results obtained by employing the Refined NHC method with $\timestep = 1~\text{fs}$.
Error bars were obtained using the Overlapping Batch Means estimator \cite{Meketon_1984, Flegal_2010} and standard uncertainty propagation formulas.

>>>>

In the previous section, we have seen that drifts in the extended energy of an NVT simulation can accumulate in the thermostat-related variables, thus preventing full-fledged instability from happening.
This allows simulations to be executed with large time steps, which could in principle be used to save computer time.
However, as we will show in this section, discretization errors can become manifest much earlier than does any risk of instability.
Fortunately, our proposed methodology is capable of mitigating the effect of these errors on the ensemble averages computed for the simulated system.

In Fig.~\ref{fig:properties}, we present some thermodynamic properties computed for the liquid water system using several NVT-MD methods with different time step sizes.
All simulations were carried out by specifying $T = 298.0~\text{K}$ and the analyzed quantities are the estimated temperature $T_\text{est} = \frac{2\langle K \rangle}{(6N-3)k_\mini{B}}$,
the mean potential energy per molecule $\langle U/N \rangle$,
the mean internal virial per molecule $\langle W/N \rangle$, and
the constant-volume specific heat $c_\mini{V} = \frac{\langle \Ham{}^2 \rangle - \langle \Ham{} \rangle^2}{N k_\mini{B} T^2}$.
The internal virial $W$ is computed as described in Ref.~\citenum{Silveira_2017}.
All symbols in Fig.~\ref{fig:properties} represent results obtained after $18~\text{ns}$ of production time, while the lines are just guides for the eyes.
The horizontal, dotted lines are meant to help visualizing the departure of the computed averages from their expected values.
They correspond to the results obtained by employing the Refined NHC method with $\timestep = 1~\text{fs}$.
Error bars were obtained using the Overlapping Batch Means estimator \cite{Meketon_1984, Flegal_2010} and standard uncertainty propagation formulas.
